---
title: "Demouge Benjamin - PW 6"
author: "Benjamin Demouge"
output: html_notebook
---


### Question 1
```{r}
iris <- read.csv("datasets/iris.data")
head(iris)
```

### Question 2
```{r}
par(mfrow = c(2, 2))
boxplot(iris$sepal_length~iris$class)
boxplot(iris$sepal_width~iris$class)
boxplot(iris$petal_length~iris$class)
boxplot(iris$petal_width~iris$class)
```

### Question 3
```{r}
library(ggplot2)

# histogram of sepal_length
ggplot(iris, aes(x = sepal_length, fill = class)) +
  geom_histogram(binwidth = .2, alpha = .5)
# histogram of sepal_width
ggplot(iris, aes(x = sepal_width, fill = class)) +
  geom_histogram(binwidth = .2, alpha = .5)
# histogram of petal_length
ggplot(iris, aes(x = petal_length, fill = class)) +
  geom_histogram(binwidth = .2, alpha = .5)
# histogram of petal_width
ggplot(iris, aes(x = petal_width, fill = class)) +
  geom_histogram(binwidth = .2, alpha = .5)
```

### Question 4
```{r}
pcairis <- princomp(iris[, -5], cor=T)
str(pcairis)
```
```{r}
summary(pcairis)
```
```{r}
plot(pcairis)
```

```{r}
biplot(pcairis)
```


### Question 5
```{r}
library(factoextra)
fviz_eig(pcairis, addlabels = TRUE)
```
```{r}
fviz_pca_var(
  pcairis,
  col.var = "cos2",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE # Évite le chevauchement de texte
)
```

```{r}
fviz_pca_ind(
  pcairis,
  col.ind = "cos2",
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE # Évite le chevauchement de texte
)
```


```

```{r}
fviz_pca_biplot(
    pcairis,
    repel = TRUE,
    col.var = "#2E9FDF", # Couleur des variables
    col.ind = "#696969"  # Couleur des individues
)
```

```{r}
fviz_contrib(pcairis, choice = "var", axes = 1)
```

```{r}
fviz_contrib(pcairis, choice = "var", axes = 2)
```


### Question 6
```{r}
X <- iris[,-5]
y <- iris[,5]
```

### Question 7
```{r}
X_scaled <- scale(X)
summary(X_scaled)
```

### Question 8
```{r}
cov_matrix <- cov(X_scaled)
cov_matrix
```

### Question 9
```{r}
eigen_decomposition <- eigen(cov_matrix)
eigen_decomposition
```
With the eigen function, you obtain the eigenvalues and the eigenvectors.

### Question 10
```{r}
corr_matrix <- cor(X_scaled)
eigen(corr_matrix)
```

### Question 11
```{r}
corr_matrix <- cor(X)
eigen(corr_matrix)
```
It is the same values with X_scaled and with X.


### Question 12

```{r}
explained_variations <- eigen_decomposition$values / sum(eigen_decomposition$values)
explained_variations
```

```{r}
explained_variations_cumulated <- cumsum(explained_variations)
explained_variations_cumulated
```

### Question 13
```{r}
plot(explained_variations, type="b")
```

### Question 14
```{r}
projection_matrix <- eigen_decomposition$vectors[, 1:2]
```


### Question 15
```{r}
Y <- X_scaled %*% projection_matrix
dim(Y)
```
```{r}
head(Y)
```

### Question 16
```{r}
plot(Y, xlab="PC1", ylab="PC2")
```

### Question 17
```
```{r}
my_cols <- c("red", "green", "blue")
plot(
  Y,
  col=my_cols[as.integer(as.factor(iris$class))],
  xlab = "PC1",
  ylab = "PC2",
  main="Plot of the data on the new feature space",
)

legend(
  2,
  2.6,
  legend=unique(iris$class),
  col=my_cols,
  lty=1:2,
  cex=0.8
)
```
