---
title: "R Notebook"
output: html_notebook
---
```{r}
dt = read.csv('iris.data')
```
2. Compare the means and the quartiles of the 3 different flower classes for the 4 different features (Plot 4 boxplots into the same figure).
```{r}
library(ggplot2)
par(mfrow=c(2,2))
boxplot(dt$sepal_length~dt$class)
boxplot(dt$sepal_width~dt$class)
boxplot(dt$petal_length~dt$class)
boxplot(dt$petal_width~dt$class)
```
3. To explore how the 3 different flower classes are distributed along the 4 different features, visualize them via histograms using the following code.
```{r}
# histogram of sepal_length
ggplot(iris, aes(x=sepal_length, fill=class)) +
  geom_histogram(binwidth=.2, alpha=.5)
# histogram of sepal_width
ggplot(iris, aes(x=sepal_width, fill=class)) +
  geom_histogram(binwidth=.2, alpha=.5)
# histogram of petal_length
ggplot(iris, aes(x=petal_length, fill=class)) +
  geom_histogram(binwidth=.2, alpha=.5)
# histogram of petal_width
ggplot(iris, aes(x=petal_width, fill=class)) +
  geom_histogram(binwidth=.2, alpha=.5)
```
4. Apply a PCA on the Iris dataset using the princomp() function and interpret the results.
```{r}

```
5. Using factoextra package plot the following:

The scree plot.
The graph of individuals.
The graph of variables.
The biplot graph.
The contributions of the variables to the first 2 principal components.
```{r}

```
6. First step, split the iris dataset into data  
X
  and class labels  
y
 .
```{r}

```
7. Scale the 4 features. Store the scaled matrix into a new one (for example, name it X_scaled).

```{r}

```
8. The classic approach to PCA is to perform the eigendecomposition on the covariance matrix  
Σ
 , which is a  
p
×
p
  matrix where each element represents the covariance between two features. Compute the Covariance Matrix of the scaled features (Print the results).
```{r}

```

9. Perform an eigen decomposition on the covariance matrix. Compute the Eigen vectors and the Eigenvalues (you can use the eigen() function). What do you obtain?
```{r}
```
10. Perform an eigendecomposition of the standardized data based on the correlation matrix.
```{r}
```
11. Perform an eigendecomposition of the raw data based on the correlation matrix. Compare the obtained results with the previous question
```{r}
```
12. Calculate the individual explained variation and the cumulative explained variation of each principal component. Show the results
```{r}
```
13. Plot the individual explained variation. (scree plot)
```{r}
```
14. Construct the projection matrix that will be used to transform the Iris data onto the new feature subspace
```{r}
```
15. Compute  
Y
  (Recall the  
Y
  is the matrix of scores,  
A
  is the matrix of loadings).
```{r}
```
16. Plot the observations on the new feature space. Name the axis PC1 and PC2.
```{r}
```
17. On the same plot, color the observations (the flowers) with respect to their flower classes.
```{r}
```


```{r}
```
